A Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-02-02 01-intro.tex
Revisão da Introdução.

\CEU~\cite{ceu.sensys13} is a programming language for embedded soft real-time
systems that aims to offer a concurrent and expressive alternative to C and C++
with the characteristics that follow:

\begin{description}
\item [Reactive:] code only executes in reactions to events.
\item [Structured:] programs use structured control mechanisms, such as
    \code{await} (to suspend a line of execution), and \code{par} (to combine
    multiple lines of execution).
\item [Synchronous:] reactions run atomically and to completion on each line of
    execution, i.e., there's no implicit preemption or real parallelism.
\end{description}

Structured reactive programming eliminates the
\emph{callback hell}~\cite{TODO}, letting programmers write code in
direct/sequential style \cite{TODO}.
%
Previous work evaluates the expressiveness of \CEU in comparison to
event-driven code in C and attests a reduction in source code size (around
25\%) with a small increase in memory usage (around 5--10\% for \emph{text} and 
\emph{data})~\cite{ceu.sensys13}.
%
\CEU has been used in the context of wireless sensor
networks~\cite{ceu.sensys13,ceu.terra},
multimedia systems~\cite{ceu.media.webmedia16}, and
games \footnote{\fs{TODO}}.
It has also been used as an alternative language in an undergradutate-level
course on embedded systems for the past 5 years.

\fs{Embedded systems requirements, determinism, termination}

\begin{itemize}
\item \CEU
    \begin{itemize}
    \item variaveis compartilhadas
    \item concorrencia segura
    \item terminacao e determinismo
    \end{itemize}
\item Modelo Sincrono
    \begin{itemize}
    \item conhecido como ammenable to verification
    \item valido para \CEU?
    \end{itemize}
\end{itemize}

%only the static semantics for embedded systems

>>>>

\CEU~\cite{ceu.sensys13,ceu.tecs17} is a Esterel-based~\cite{esterel.ieee91}
programming language for embedded soft real-time systems that aims to offer a
concurrent, safe, and expressive alternative to C with the characteristics that
follow:
%
\begin{description}
\item [Reactive:] code only executes in reactions to events.
\item [Structured:] programs use structured control mechanisms, such as
    \code{await} (to suspend a line of execution), and \code{par} (to combine
    multiple lines of execution).
\item [Synchronous:] reactions run atomically and to completion on each line of
    execution, i.e., there's no implicit preemption or real parallelism.
\end{description}
%
Structured reactive programming let developers write code in direct/sequential
style, recovering from the inversion of control imposed by event-driven
execution~\cite{rp.deprecating,rp.rescala,sync_async.cooperative}.
%
Synchronous languages offer a simple run-to-completion model of execution that
enable deterministic execution and make formal reasoning tractable.
For this reason, it has been successfully adopted in safety-critical real-time
embedded systems.~\cite{rp.twelve}

Previous work in the context of embedded sensor networks evaluates the
expressiveness of \CEU in comparison to event-driven code in C and attests a
reduction in source code size (around 25\%) with a small increase in memory
usage (around 5--10\% for \emph{text} and \emph{data})~\cite{ceu.sensys13}.
%
\CEU has also been used in the context of multimedia
systems~\cite{ceumedia.webmedia16} and games~\cite{ceu.mod15}, and as an
alternative language in an undergraduate-level course on embedded systems for
the past 6 years.

\CEU inherits the synchronous and imperative mindset of Esterel but adopts a
simpler semantics with fine-grained execution control.~\cite{ceu.tecs17}
%
The list that follows summarizes the semantic peculiarities of \CEU:
%
\begin{itemize}
    \item Stack-based execution for internal events, which provides a limited
          form of coroutines.
    \item Fine-grained, intra-reaction deterministic execution, which allows
          programs to safely share memory.
    \item Finalization mechanism for abortion of lines of execution, which
          safely release external resources.
    \item First-class synchronized timers.
\end{itemize}

In this work, we present a formal semantics for a subset of \CEU that focus on
its peculiarities in comparison to other synchronous languages.
\begin{itemize}
    \item qual a abordagem / operational semantics / dois passos
    \item quais os resultados / provas
    \item quais os desafios e limitações
    \item \gl{TODO}
\end{itemize}

\fs{Descrever seções.}

### 2018-02-02 index.tex
Revisão da Introdução.

\begin{abstract}
\CEU is a synchronous programming language for embedded soft real-time systems.
%
It focus on control-flow safety features, such as safe shared-memory
concurrency and safe abortion of lines of execution, while enforcing internal
determinism and time-bounded reactions to the environment.
%
In this work, we present a small-step structural operational semantics for
\CEU and a proof that reactions are deterministic and always terminate:
%
For a given timeline of input events, multiple executions of the same program
always react in bounded time and arrive at the same final memory state.
%
%We also discuss some equivalence results and the relation between the proposed
%semantics and its actual implementation.
\end{abstract}

>>>>

\begin{abstract}
\CEU is a synchronous programming language for embedded soft real-time systems.
%
It focus on control-flow safety features, such as safe shared-memory
concurrency and safe abortion of lines of execution, while enforcing internal
determinism and time-bounded reactions to the environment.
%
In this work, we present a small-step structural operational semantics for
\CEU and a proof that reactions are deterministic and always terminate:
%
For a given arbitrary timeline of input events, multiple executions of the same
program always react in bounded time and arrive at the same final memory state.
%
%We also discuss some equivalence results and the relation between the proposed
%semantics and its actual implementation.
\end{abstract}
